Making Sequential Programs Deterministic

### 2019-04-06 design.tex
Remove duplication
Tweaks in 2.1-2.2

This section presents the key aspects of our type system,
called \ourTypeSystem, in the context of a core calculus for an
object-oriented language.

@@

In a type system,
a \textit{type} represents a set of values.  It abstracts or restricts the
set of possible run-time values that an expression may evaluate to.
A programming language provides \emph{basetypes}, such as \<Int>.
A \textit{type qualifier} on a basetype adds additional constraints;
that is, it reduces the size of the set of values.
An example type qualifier is \<Positive>, and a type (which combines a qualifier
and a basetype) is \<Positive Int>.
A polymorphic type abstraction such as \<List> can be instantiated by a type argument,
as in \ptype{List}{Positive Int}.

@@

A type checker verifies the types written in a program.
\OurTypeSystem checks all the standard typing rules including the ones
shown in \cref{typecheck-rules-standard}.  For \ourTypeSystem, $\tau$
represents a qualified type rather than a basetype.

A type qualifier constrains the set of possible run-time values, that is,
$\kappa \ \beta \sqsubseteq \beta$.
As a result, a qualifier type system does not allow any values that the
original type system does not.
However, the qualifier type system may reject more programs, and thus
affords stronger guarantees.
% (Our implementation \theDeterminismChecker runs the underlying base type
% system, then issues additional errors and warnings.)

Type qualifier systems are sometimes defined independently of the
underlying type system.  Because there are interactions between the
basetypes and the type qualifiers, defining them together improves
precision, which is important in practice.

@@

\Cref{typecheck-rules-standard} shows a sample of standard typing rules for
an object-oriented programming language.

@@

\<OrderNonDet> may only be written on collections and maps.
A map is a dictionary or associative array, such as a hash table.
\OurTypeSystem largely treats a map as a collection of key--value pairs.
Both collections and maps may be \<Det>, \<OrderNonDet>, or \<NonDet>, and
the element types can be specified independently of the collection type,
including qualifiers.

>>>>

This section presents the key aspects of our type system
% , called \ourTypeSystem,
in the context of a core calculus for an
object-oriented language.

@@

One way to view a type is as
a set of values.  A type abstracts or restricts the
set of possible run-time values that an expression may evaluate to
and the operations that may be performed.
A programming language provides \emph{basetypes}, such as \<Int>.
A \textit{type qualifier} on a basetype adds additional constraints;
that is, it reduces the size of the set of values.
An example type qualifier is \<Positive>, and a type (which combines a qualifier
and a basetype) is \<Positive Int>.
A polymorphic type abstraction such as \<List> can be instantiated by a type argument,
as in \ptype{List}{Positive Int}.

@@

A type checker verifies the types written in a program.
\OurTypeSystem checks all the standard typing rules including the ones
shown in \cref{typecheck-rules-standard}, which
shows a sample of standard typing rules for
an object-oriented programming language.
% For \ourTypeSystem, $\tau$
% represents a qualified type rather than a basetype.

A type qualifier constrains the set of possible run-time values, that is,
$\kappa \ \beta \sqsubseteq \beta$.
As a result, a qualifier type system does not allow any values that the
original type system does not, in the same program without qualifiers.
However, the qualifier type system may reject more programs, and thus
affords stronger guarantees.
% (Our implementation \theDeterminismChecker runs the underlying base type
% system, then issues additional errors and warnings.)

Type qualifier systems are sometimes defined independently of the
underlying type system.  In \ourTypeSystem, there are interactions between the
basetypes and the type qualifiers.  Defining them together improves
precision, which is important in practice.

@@

@@

\<OrderNonDet> may only be written on collections and maps.
A map is a dictionary or associative array, such as a hash table.
\OurTypeSystem largely treats a map as a collection of key--value pairs.
Both collections and maps may be \<Det>, \<OrderNonDet>, or \<NonDet>, and
the element types can be specified independently of the collection type,
including their qualifiers.
